---
title: "Reproducible document"
author: "Nemanja Stanojevic"
format: html
---
## Basics of R
```{r}
```

```{r}
1 + 1
```



## About me
-  Nemanja
-  Stanojevic
-  Roskilde University
I am from *Serbia*, I am a **Medical doctor**

## Simple code
```{r}
3 * 3
```
## Testing for Git
The weather is awful
  
```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)
small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```

### Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
@fig-bmi-histo


```{r}
#| fig-cap: "Glykemi"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Putting two plots side by side

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by-side}
#| label: "fig-bmi-glycemia"
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|  - "Distribution of BMI."
#|  - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose."
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```


## Pushing changes to remote

## Pulling changes locally

## Selecting data

```{r}
select(post_meal_data, Age)
```
```{r}
select(post_meal_data, Age, BMI, Weight)
```
```{r}
select(post_meal_data, -Group)
```

```{r}
select(post_meal_data, starts_with("PG"))
```

```{r}
select(post_meal_data, ends_with("DXA"))
```
```{r}
select(post_meal_data, contains("0"))
```
## Renaming specific columns


```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```


## Select and rename tasks


```{r}
select(post_meal_data, contains("OGTT"))
```

```{r}
select(post_meal_data, contains(".."), -contains("..."))
```
```{r}
post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```


  
